---
title: "spencer-dev-notebook"
author: "Spencer"
format: html
editor: visual
---

Questions:

- For any functions with multiple output types (e.g., `em_loadgenerator()`) how should their outputs be organized? A list? Currently `em_buildgenerator` returns a list with 2 objects, a `landscape` and a `generator`. The list itself is a `generator` which I don't think makes sense.
- Where does `calc_step` go? `functions.R` or `generate-examples.R`? I don't know how to untangle having a movement model function that doesn't require some version of `calc_step` to successfully build a generator.
- When applying the toy low/mid/high movement model to a landscape you need to use `dplyr` functions (mutate and case_when) so it loses its class. Currently the quick fix is to redefine it as a `landscape` class after working with it but a permanent option is to set up methods for mutate & case_when.

To do:

- In `eigenmove-vignette` either use `rescale_landscape()` or code in assigning qualities to intensity values of pixels. Assigning qualities is no longer included in `em_loadlandscape()`.

#### Aug 8 2026

Started `building-functions` branch to work on coding eigenmove functions according to what's written in the shared canva whiteboard and zulip correspondence. Will borrow what's possible from existing functions, looking first to Eric's updated code from the `metapopulations-and-movement-models` repository.

#### Aug 10 2026

Added `em_loadlandscape` as a function to canva whiteboard. Transferred canva table to a .qmd in this repo: eigenmove-structure.qmd.

Added `em_loadlandscape` to functions.R. Problem is that we are loading a landscape and assigning qualities in generate-examples.R.

Solution is to just code in the example vignette habitat qualities.

```{r}
a <- em_loadlandscape("../dev/gp-example.png", keep_channels = TRUE)
head(a)
```

So `em_loadlandscape()` is working with and without keep channels. generate-landscape-vignette.Rmd is still fine because it never used loadlandscape() in the first place. When I write eigenmove-vignette.Rmd I will probably have to just hardcode in assigning qualities to intensity values (or maybe use that rescale_landscape() function?).

We want a function `em_randomwalk` to be a functional that builds a discrete movement model (object type `movemodel`) based on step_length, step_speed, and preference. We also want a function `em_movemodel` that describes movement on a landscape (object type `generator`) in discrete or continuous time, taking `TBD` inputs.

Currently we have `em_create_example_Q` which ...

#### Aug 11 2026

Asked claude LLM for help separating out em_randomwalk from em_create_example_Q to make a function that is not tied to a landscape.

```{r}
a_movemodel <- em_randomwalk()
```

So now I have a landscape `a` and a movement model `a_movemodel`, now we want to apply the model to the landscape and build a generator matrix. First, we need to assign 3 habitat types for the `movemodel` to read. Built into `em_randomwalk()` are different preferences for 3 habitat types, but our landscape only has pixel intensity, not "type" yet, so I will assign different intensity values to different types.

```{r}
a = a |> 
  dplyr::mutate(type = dplyr::case_when(intensity < 0.5 ~ "high", 
                                 intensity ==  1 ~ "low",
                                 TRUE ~ "mid"),
                type = factor(type, levels = c("low", "mid", "high")))
class(a) <- c("landscape", class(a))
a_generator <- em_buildgenerator(landscape = a, movemodel = a_movemodel)
summary(a_generator)
```

Need to make `landscape` and `generator` both classes. Need to check that `em_buildgenerator` is working correctly. Need to make sure the generator is most efficiently tied to the landscape.

#### Aug 12 2026

Now the return from `em_loadlandscape` is a `landscape` class. It's a little weird if you want to apply an example random walk on it because you need to redefine it as a class after (since you need to use mutate & case_when). Currently the quick fix is to redefine it as a `landscape` class after working with it but a permanent option is to set up methods for mutate & case_when.
