---
title: "spencer-dev-notebook"
author: "Spencer"
format: html
editor: visual
---

#### Checks & Todos

`probmove` is commented out in functions.R.\
Need to fix vignettes to work with new landscape.R and `calc_step()` (where is it? see below.)

#### Questions and Problems:

- Where does `calc_step` go? `functions.R` or `generate-examples.R`? Doesn't a movement model function require some version of `calc_step` to successfully build a `generator`? Is the creation of `calc_step` supposed to be included in `em_randomwalk()`?
- When applying the toy low/mid/high movement model to a landscape you need to use `dplyr` functions (mutate and case_when) so it loses its class. Currently the quick fix is to redefine it as a `landscape` class after working with it but a permanent option is to set up methods for mutate & case_when.
- We have `em_buildgenerator()` to apply a movement model to a landscape, then we have `em_loadgenerator()` which is supposed to apply a pre-existing generator to a landscape, but isn't a generator by definition a movement model applied to a landscape?
- In the planned Objects table from the canva we have `paths` as a tibble. I coded it in `em_simmove()` as a list because that's what made sense to me. I will re-write the code to produce it as a tibble, but I can't picture how it would look.

To do:

- In `eigenmove-vignette` either use `rescale_landscape()` or code in assigning qualities to intensity values of pixels. Assigning qualities is no longer included in `em_loadlandscape()`.
- Add functionality to `em_simmove()` to accept an `eigenmove` object as well as a generator.
- All dependencies are currently imports. Need to sort out any possible "suggests". Maybe igraph? sf?

#### Aug 08 2026

Started `building-functions` branch to work on coding eigenmove functions according to what's written in the shared canva whiteboard and zulip correspondence. Will borrow what's possible from existing functions, looking first to Eric's updated code from the `metapopulations-and-movement-models` repository.

#### Aug 10 2026

Added `em_loadlandscape` as a function to canva whiteboard. Transferred canva table to a .qmd in this repo: eigenmove-structure.qmd.

Added `em_loadlandscape` to functions.R. Problem is that we are loading a landscape and assigning qualities in generate-examples.R.

Solution is to just code in the example vignette habitat qualities.

```{r}
library(devtools)
a <- em_loadlandscape("../dev/gp-example.png", keep_channels = TRUE)
head(a)
```

So `em_loadlandscape()` is working with and without keep channels. generate-landscape-vignette.Rmd is still fine because it never used loadlandscape() in the first place. When I write eigenmove-vignette.Rmd I will probably have to just hardcode in assigning qualities to intensity values (or maybe use that rescale_landscape() function?).

We want a function `em_randomwalk` to be a functional that builds a discrete movement model (object type `movemodel`) based on step_length, step_speed, and preference. We also want a function `em_movemodel` that describes movement on a landscape (object type `generator`) in discrete or continuous time, taking `TBD` inputs.

Currently we have `em_create_example_Q` which ...

#### Aug 11 2026

Asked claude LLM for help separating out em_randomwalk from em_create_example_Q to make a function that is not tied to a landscape.

```{r}
a_movemodel <- em_randomwalk()
```

So now I have a landscape `a` and a movement model `a_movemodel`, now we want to apply the model to the landscape and build a generator matrix. First, we need to assign 3 habitat types for the `movemodel` to read. Built into `em_randomwalk()` are different preferences for 3 habitat types, but our landscape only has pixel intensity, not "type" yet, so I will assign different intensity values to different types.

```{r}
a = a |> 
  dplyr::mutate(type = dplyr::case_when(intensity < 0.5 ~ "high", 
                                 intensity ==  1 ~ "low",
                                 TRUE ~ "mid"),
                type = factor(type, levels = c("low", "mid", "high")))
class(a) <- c("landscape", class(a))
a_generator <- em_buildgenerator(landscape = a, movemodel = a_movemodel)
summary(a_generator)
```

Need to make `landscape` and `generator` both classes. Need to check that `em_buildgenerator` is working correctly. Need to make sure the generator is most efficiently tied to the landscape.

#### Aug 12 2026

Now the return from `em_loadlandscape` is a `landscape` class. It's a little weird if you want to apply an example random walk on it because you need to redefine it as a class after (since you need to use mutate & case_when). Currently the quick fix is to redefine it as a `landscape` class after working with it but a permanent option is to set up methods for mutate & case_when.

#### Aug 31 2026

Return after couple weeks focusing on reading instead. Had success during 2nd week of August and set aside to focus on reading.

Next up, `em_simmove()`. It will return `paths` which is cleanest to make a list that describes sets of steps taken on a `landscape`. Arguments:\
\
-`generator`​ (list including a `landscape`​ and coupled `generator`​ movement matrix)\
-`steps`​ (a parameter of number of random steps each path contains)\
-`replicates`​ (a parameter of number of repeated samplings of random steps for each path)\
-`origins`​ (a parameter of number of different paths to take, each from a different random x,y origin location)

Each paths entry will need to have an origin location drawn randomly from `landscape`​. Each origin will have `replicate`​ number of randomly generated sets of consecutive `steps`​ number of steps, drawn from the stochastic matrix `generator`​. Later, each `paths`​ entry will be visualized ontop of the landscape.

```{r}
a_paths <- em_simmove(a_generator, 
                      steps = 5,
                      replicates = 3,
                      origins = 2)
```

Created separate scripts defining R6 classes for landscapes and generators. Need to rewrite these functions to implement the content from eigenmove-structure.qmd.

#### Sep 01 2026

Working on landscape.R.

Looking good so far. I created a test script to test to make sure it's all running correctly. Haven't tested it yet, need to understand the testthat workflow. Some notes: `probmove` is commented out in functions.R. There will probably be an error with `em_loadlandscape()` with the line that flips the y coordinates. Not sure if using a global variable `y` here will work.

#### Sep 10 2026

Maybe fixed the line that flips the y variable by adding .data\$y which accesses rlang. Also added `Importfrom` s to Roxygen comments.
