---
title: "generate-landscape-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{generate-landscape-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eigenmove)
```

In this vignette you will see a complete example of eigenmove's capabilities. We will show you how to generate a toy landscape using Gaussian Processes (`create_GP_landscape()`) and how to compute random walks on that landscape according to a simple movement model (`load_landscape()`, `calc_step()`, `rescale_landscape()`). This will result in a *dispersal matrix* which can then be processed through the core `eigenmove` functionality of estimating patches on a landscape. If you already have a dispersal matrix and just want to skip to estimating patches, see `vignette("eigenmove-vignette")`.

### Creating a toy landscape

Use `create_GP_landscape()` to simulate a toy landscape via a Gaussian process for use with simple random walk movement model. 

```{r}
small_landscape <- create_GP_landscape(
  landscape_width = 32,
  landscape_height = 32,
  patch_scale = 10
)
```

Run create_GP_landscape to generate a data frame. e.g. current_GP \<- create_GP_landscape() 2 --- Run rescale_landscape using the value column of the create_GP_landscape to generate a new column with discrete low, mid, high values. e.g. current_GP$type <- rescale_landscape(current_GP$value)

  #This function randomly generates a new landscape using what's called a
  #Gaussian process; this assumes that the random value at each point in the
  #landscape follows a normal distribution, but that the random values for
  #points close to one another are correlated, so they have similar values.
  #Probably the best intro to GPs is here:
  #https://distill.pub/2019/visual-exploration-gaussian-processes/ although it
  #focuses more on using GPs for fitting data than for simulating data

  #landscape_width: specifies the width of the landscape in pixels
  #landscape_height: specifies the height of the landscape in pixels
  #patch_scale: specifies the size of the patch. Larger values of patch_scale
  #correspond to higher correlations between distant points, so larger (but less
  #common) patches
